---
title: "Clinical Characteristics stat"
output: html_document
date: "2025-01-09"
---

__Load library__ 

```{r}
library(gplots)
```

__Comparisons of categorical data between diabetic and controls were performed using Chi-square or Fisher's test followed by corrections of Yates and Bonferroni.__

We show sex and comorbidities as example

Import sex data and converte dataframe in table

```{r}
sex <- read.table("../data/Sex.csv", row.names=1, header = TRUE,
                  sep=";", stringsAsFactors=TRUE)
sex_dt <- as.table(as.matrix(sex))
sex_dt

```

__Plot the data__
```{r}
balloonplot(t(sex_dt), main = "Sex", xlab = "", ylab = "",
            label = TRUE, show.margins = FALSE, dotsize = 15)
```


__Chi-square test with Yates correction__
```{r}
chisq.test(sex_dt, correct = TRUE)
```
Import comorbidities data and converte dataframe in table
```{r}
Comorbidities <- read.table("../data/Comorbidities.csv", row.names=1, header = TRUE,
                     sep=";", stringsAsFactors=TRUE)
Comorbidities_dt <- as.table(as.matrix(Comorbidities))
Comorbidities_dt
```

__Plot the data__
```{r}
balloonplot(t(Comorbidities_dt), main ="Comorbidities", xlab ="", ylab="",
            label = TRUE, show.margins = FALSE, dotsize=15)
```

__Fisher's test with Bonferroni corrections__

```{r}
fisher_result <- fisher.test(sex_dt)
fisher_result

p_values <- c(fisher_result$p.value)
adjusted_p_values <- p.adjust(p_values, method = "bonferroni")
adjusted_p_values
```

# Age comparison between groups

__Load library__ 
```{r}
library(readxl)
library(tidyverse)
library(introdataviz)
library(ggpubr)
```

Import data
```{r}
AgeStat <- read_excel("../data/Data_EpidemiologicalCharacteristics.xlsx", sheet = 3)
AgeStat
```

__Statistical Analysis__
```{r}
# Step 1: Decide whether to use the Shapiro-Wilk or Kolmogorov-Smirnov normality test
decidir_prueba_normalidad <- function(x) {
  if (length(x) < 50) {
    return(shapiro.test(x)$p.value)
  } else {
    return(ks.test(scale(x), "pnorm")$p.value)
  }
}

# Step 2: Apply the normality test function for each group
normality_test <- AgeStat %>%
  group_by(Group) %>%
  summarise(p_value = decidir_prueba_normalidad(Age))

# Step 3: Check if all groups are normal (p-value > 0.05)
if (all(normality_test$p_value > 0.05)) {
  test_type <- "t.test"
} else {
  test_type <- "wilcox.test"
}

```

__Plot__
```{r}
# Define color and comparisons
colours <- c("#0000CD", "#FF3030")
my_comparisons <- list(
  c("Diabetic", "NonDiabetic"))
rain_height <- 0.1

ggplot(AgeStat, aes(x = Group, y = Age, fill = Group)) +
  geom_flat_violin(trim = FALSE, alpha = 0.4, position = position_nudge(x = rain_height + .05)) +
  geom_point(aes(colour = Group), size = 2, alpha = .5, show.legend = FALSE, position = position_jitter(width = rain_height, height = 0)) +
  geom_boxplot(width = rain_height, alpha = 0.4, show.legend = FALSE, outlier.shape = NA, position = position_nudge(x = -rain_height * 2)) +
  stat_summary(fun.data = mean_cl_normal, mapping = aes(color = Group), show.legend = FALSE, position = position_nudge(x = rain_height * 3)) +
  scale_fill_manual(values = colours, name = "") +
  scale_colour_manual(values = colours) +
  theme_classic() +
  theme(legend.position = "none") +
  labs(x = "Group", y = "Age", title = paste("Test:", ifelse(test_type == "t.test", "T-test", "Mann-Whitney U"))) +
  stat_compare_means(method = test_type, comparisons = my_comparisons, label = "p.format")
```

